The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-06-22 WavesData-2019/LatexNotes/notes.tex
NM

% New
\newcommand{\Hp}{H^{+}}
\newcommand{\Hm}{H^{-}}
\newcommand{\Gp}{\Gibbs^{+}}
\newcommand{\Gm}{\Gibbs^{-}}
\newcommand{\meanp}[1]{\mean{#1}_{+}}
\newcommand{\meanm}[1]{\mean{#1}_{-}}

@@

In \eqref{ndKdV}, position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$. Further, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
Then the Hamiltonian is given by
\begin{align}
\label{Hamiltonian}
& \Hm = \ampp H_3 - \mu H_2 \\
& \Hp = \ampp D_0^{-3/2} H_3 - \mu D_0^{1/2} H_2
\end{align}
where
\begin{align}
&H_3 = \frac{1}{6} \intt u^3 \dx	\, , \\
&H_2 = \frac{1}{2} \intt u_x^2 \dx	\, .
\end{align}

@@

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table of parameters}
\label{paramtable}
\begin{tabular}{l l l l}
\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
\hline Description & Notation and definition & Units & Value in experiments \\
\hline
Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2}$		& length		& 0.03--0.3 cm \\
Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
Amplitude-to-depth ratio		& $\eps = a / h_0$				& dimensionless	& 0.0024--0.024 \\
Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
Amplitude parameter			& $\ampp = \eps \delta^{-2} = ag h_0^{-2} f_p^{-2}$	
& dimensionless		& 0.05--0.5\\
Catch-all parameter		& $\mu = 4 \pi^2 / (9 \lamfac^2)$			& dimensionless		& 0.27 if $\lamfac=4$ \\
\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
\hline Description & Notation and definition & Units \\
\hline
Local depth			& $d$					& length \\
Local wavespeed		& $c = \sqrt{gd}$			& speed \\
Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
Characteristic wavelength	& $\lambda_c = \lamfac \lambda_p$	& length \\
Depth ratio			& $D = d/h_0$				& dimensionless
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

The statistical matching condition is
\begin{equation}
\meanm{\Hp} = \meanp{\Hp}
\end{equation}
An experimental observation is that the incoming skewness is small, thus
\begin{equation}
\meanm{H_3} \approx 0
\end{equation}
which immediately gives
\begin{equation}
\ampp D_0^{-3/2} \meanp{H_3} = \mu D_0^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
\end{equation}
or
\begin{equation}
\meanp{H_3} = \frac{\mu D_0^{2}}{\ampp} \left( \meanp{H_2} - \meanm{H_2} \right)
\end{equation}

Converting to dimensional (i.e.~experimental) values gives
\begin{equation}
\frac{\text{skew}(\eta)}{\Delta \mean{\eta_x^2}} =
\frac{3 \mu h_0^2 f_p^2 D_0^2 \lambda^2}{g a^3}
\end{equation}

>>>>

% New
\newcommand{\Hp}{H^{+}}
\newcommand{\Hm}{H^{-}}
\newcommand{\Gp}{\Gibbs^{+}}
\newcommand{\Gm}{\Gibbs^{-}}
\newcommand{\meanp}[1]{\mean{#1}_{+}}
\newcommand{\meanm}[1]{\mean{#1}_{-}}
\newcommand{\skw}{\text{skew}}
\newcommand{\var}{\text{var}}

@@

In \eqref{ndKdV}, position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$. Further, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
Then the Hamiltonian is given by
\begin{align}
\label{Hamiltonian}
& \Hm = \ampp H_3 - \mu H_2 \\
& \Hp = \ampp D^{-3/2} H_3 - \mu D^{1/2} H_2
\end{align}
where
\begin{align}
&H_3 = \frac{1}{6} \intt u^3 \dx	\, , \\
&H_2 = \frac{1}{2} \intt u_x^2 \dx	\, .
\end{align}

@@

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table of parameters}
\label{paramtable}
\begin{tabular}{l l l l}
\hline \multicolumn{4} { c }{Parameters that are constant throughout the domain} \\
\hline Description & Notation and definition & Units & Value in experiments \\
\hline
Peak forcing frequency		& $f_p$						& frequency	& 2 Hz \\
Characteristic wave amplitude	& $a = \mean{\eta^2}^{1/2} = \etastd$		& length		& 0.03--0.3 cm \\
Upstream wavelength			& $\lambda_0 = \sqrt{g h_0}/f_p$	& length		& 56 cm \\
Amplitude-to-depth ratio		& $\eps = a / h_0$				& dimensionless	& 0.0024--0.024 \\
Depth-to-wavelength ratio		& $\delta = h_0 / \lambda_0$		& dimensionless		& 0.22 \\
Amplitude parameter			& $\ampp = \eps \delta^{-2} = a \lambda_0^2 h_0^{-3} = ag h_0^{-2} f_p^{-2}$	
& dimensionless		& 0.05--0.5\\
Catch-all parameter		& $\mu = 4 \pi^2 / (9 \lamfac^2)$			& dimensionless		& 0.27 if $\lamfac=4$ \\
\hline \multicolumn{4} { c }{Parameters that change value across ADC} \\
\hline Description & Notation and definition & Units \\
\hline
Local depth			& $d$					& length \\
Local wavespeed		& $c = \sqrt{gd}$			& speed \\
Local peak wavelength	& $\lambda_p = \sqrt{gd}/f_p$	& length \\
Characteristic wavelength	& $\lambda_c = \lamfac \lambda_p$	& length \\
Depth ratio			& $D = d/h_0$				& dimensionless
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

The statistical matching condition is
\begin{equation}
\meanm{\Hp} = \meanp{\Hp}
\end{equation}
An experimental observation is that the incoming skewness is small, thus
\begin{equation}
\meanm{H_3} \approx 0
\end{equation}
which immediately gives
\begin{equation}
\ampp D^{-3/2} \meanp{H_3} = \mu D^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
\end{equation}
or equivalently
\begin{equation}
\meanp{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
\end{equation}
where $\Delta \mean{H_2} =  \meanp{H_2} - \meanm{H_2}$  indicates the difference between the upstream and downstream values.

To convert to dimensional, i.e.~experimental, values we use
\begin{align}
& \intt u^3 \dx = \frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
& \intt u_x^2 \dx = \left(\frac{\lambda_0}{a} \right)^2 \mean{\eta_x^2} 
= \left(\frac{\lambda_0}{a} \right)^2 \var(\eta_x)
\end{align}
Then we have the relationship
\begin{equation}
\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
